Near and Mid-infrared properties of known $z\geq5$ Quasars

### 2019-05-28 tex/DistanceModulus.tex
tex/DistanceModulus.tex added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                           
%%   The `compliation' paper for all z>5 spec-z quasars, here called "Very High Redshift Quasars" (VHzQs) and 
%%   looking into their IR properties. Motivations include putting a list of all the VHzQ in one place, looking for 
%%   signs of super-critical accretion and having a super-set of objects and their given parameters ahead of 
%%   JWST GO Cycle 1. 
%%									    
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[usenatbib]{mnras}

\input {format}

\begin{document}

\title[Very high-$z$ Quasars]
        {Near and Mid-infrared properties of known $z\geq5$ Quasars}
\author[Ross \& Cross]
       {Nicholas P. Ross$^{1}$\thanks{Corresponding Author: npross@roe.ac.uk} and Nicholas J. G. Cross$^{1}$
\\ 
$^1$Institute for Astronomy, University of Edinburgh, Royal Observatory, Edinburgh, EH9 3HJ, United Kingdom\\
}

\maketitle
\begin{abstract}
In this paper, we, for the first time since the discovery of $z\geq5$
quasars, assemble all spectroscopically confirmed very high redshift
quasars in one catalogue.  In particular we present the near
($zZyYJHK_{s}$ and $K$) infrared and mid-infrared (WISE) properties of
all 463 spectroscopically confirmed redshift $z\geq5.00$ quasars.
Using archival public WFCAM/UKIRT and VIRCAM/VISTA data we check for
photometric variability in the near-infrared that might be expected
from Super-Eddington accretion and find {\it blah}.  We present a
comprehensive series of colour-redshift and colour-colour plots and
make inferences into the hot dust properties of the very high-redshift
quasar population. Extrapolating the known quasar luminosity function
we suggest that $x$\% of the possibly detected $z\geq5$ quasars in the
current datasets have been discovered. 
All the data, analysis codes and plots used and generated here can be found at:
\href{https://github.com/d80b2t/VHzQ}{\tt github.com/d80b2t/VHzQ}.
\end{abstract}

\begin{keywords}
Astronomical data bases: surveys -- 
Quasars: general -- 
galaxies: evolution -- 
galaxies: infrared.
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Very high redshift quasars (VH$z$Q; defined here to have redshifts $z\geq5.00$) are excellent probes of the early Universe. This includes studies of the Epoch of Reionization for hydrogen \citep[see e.g.][for reviews]{Fan2006review, Mortlock2016}, the formation and build-up of supermassive black holes \citep[e.g., ][]{Rees1984, WyitheLoeb2003, Volonteri2010, Agarwal2016, Valiante2018, Latif2018, Wise2019} and early metal enrichment \citep[see e.g., ][]{Simcoe2012, Chen2017, Bosman2017}.

Super-critical accretion, where $\dot{M} > \dot{M}_{\rm Edd}$, is a viable mechanism to explain the high, potentially super-Eddington, luminosity and rapid growth of supermassive black holes in the early universe \citep[e.g.,][]{AlexanderNatarajan2014, MadauHaardtDotti2014, Volonteri2015, Pezzulli2016, Lupi2016, Pezzulli2017, Takeo2018}. Thus, one could expect VH$z$Qs to potentially vary in luminosity as they go through phases of super-critical accretion. These signatures of photometric variability should be looked for, noting the rest-frame optical emission is redshifted into the observed near-infrared (NIR) at redshifts $z>5$. Fortunately, data are now in place from deep, wide-field NIR instruments and surveys such as the Wide Field Camera (WFCAM) instrument on the United Kingdom Infra-Red Telescope (UKIRT) in the Northern Hemisphere and the VISTA InfraRed CAMera (VIRCAM) on the Visible and Infrared Survey Telescope for Astronomy (VISTA) in the Southern Hemisphere, that are necessary for identifying VH$z$Qs.

Quasars are known to be prodigious emitters of infrared emission, thought to be from the thermal emission of dust grains heated by continuum emission from the accretion disc \citep[e.g.,][]{Richards2006b, Leipski2014, Hill2014, Hickox2017}. Observations in the mid-infrared, e.g. $\sim$3-30$\mu$m allow discrimination between AGN\footnote{Historically, ``quasars'' and ``Active Galactic Nuclei (AGN)'' have described different
luminosity/classes of objects. In recognition of the fact that both terms describe accreting supermassive black holes, we use these terms interchangeably, with a preference for quasar, since we are generally in the higher-$L$ regime \citep[e.g.][]{Haardt2016book}.}  and passive galaxies due to the 1.6$\mu$m ``bump'' entering the MIR at $z\approx0.8-0.9$ \citep[e.g., ][]{Wright1994, Sawicki2002, Lacy2004, Stern2005, Richards2006b, Timlin2016} as well as between AGN and star-forming galaxies due to the presence of Polycyclic Aromatic Hydrocarbon (PAHs) at $\lambda
>3\mu$m \citep[e.g., ][]{Yan2007, Tielens2008}.

\citet{Jiang2006dust} and \citet{Jiang2010} report on the discovery of
a quasar without hot-dust emission in a sample of 21 $z\approx6$
quasars. Such apparently hot-dust-free quasars have no counterparts at
low redshift. Moreover, those authors demonstrate that the hot-dust
abundance in the 21 quasars builds up in tandem with the growth of the
central black hole. But understanding how dust first forms and appears
in the central engine remains an open question \citep{WangR2008,
WangR2011}.

WISE mapped the sky in 4 passbands, in bands centered at wavelengths
of 3.4, 4.6, 12, and 23$\mu$m. The all sky `ALLWISE' catalogue
release, contains nearly 750 million detections at
high-significance\footnote{\href{wise2.ipac.caltech.edu/docs/release/allwise/expsup/sec2\_1.html}{wise2.ipac.caltech.edu/docs/release/allwise/expsup/sec2\_1.html}},
of which over 4.5M AGN candidates have been identified with 90\%
reliability \citep{Assef2018}.  \citet{Blain2013} presented WISE
mid-infrared (MIR) detections of 17 (55\%) of the then known 31
quasars at $z > 6$. However, \citet{Blain2013} was compiled with the
WISE `All-Sky' data release, as opposed to the superior AllWISE
catalogues. That sample only examined the 31 known $z>6$ quasars; our
sample has 170 objects with redshift $z \geq 6.00$ (with $x$ detected
in WISE {\bf NPR to double check!!}). \citet{Banados2016} reports WISE
W1, W2, W3 and W4 magnitudes for the Panoramic Survey Telescope and
Rapid Response System 1 \citep[Pan-STARRS1, PS1;][]{Kaiser2002,
Kaiser2010}, but with no further investigation into the reddest WISE
waveband for the VH$z$Qs.

Critically, we now have available to us new W1 and W2 photometry from
the `unWISE Source Catalog' \citep[][]{Schlafly_Meisner2018}, a
WISE-selected catalogue that is based on significantly deeper imaging
and has a more extensive modeling of crowded regions than the ALLWISE
release. For the first time in a catalogue, unWISE takes advantage of
the ongoing mid-IR Near-Earth Object Wide-Field Infrared Survey
Explorer Reactivation mission \citep[NEOWISE-R; ][]{Mainzer2014}, and
achieves depths $\sim$0.7 mag deeper than ALLWISE (in W1/2).  This
additional depth is a significant advantage in the detection and study
of VH$z$Qs in the 3-5 micron regime.

Here we present for the first time the combined near-infrared properites (from UKIRT and VIRCAM) and the new mid-infrared unWISE for all the spectroscopically known $z\geq5.00$ quasars. Our motivations are numerous and include: {\it (i)} establishing the first complete catalogue of $z>5.00$ quasars since the pioneering work from SDSS; {\it (ii)} utilizing all the WFCAM and VISTA near-infrared photometry available for the quasars; {\it (iii)} making the first study of near- and mid-IR variability of the VHzQ population and {\it (iv)} establishing the photometric properties for upcoming surveys and telescopes including the Large Synoptic Survey Telescope (LSST)\footnote{\href{https://www.lsst.org}{{lsst.org}}}, ESA {\it Euclid}\footnote{\href{https://sci.esa.int/euclid/}{sci.esa.int/euclid/}} and the {\it James Webb Space Telescope} (JWST)\footnote{\href{https://www.jwst.nasa.gov/}{jwst.nasa.gov};}$^,$\footnote{\href{https://sci.esa.int/jwst/}{sci.esa.int/jwst};}$^,$\footnote{\href{https://www.asc-csa.gc.ca/eng/satellites/jwst/}{www.asc-csa.gc.ca/eng/satellites/jwst};}$^,$\footnote{\href{https://jwst.stsci.edu/}{jwst.stsci.edu}}. We chose redshift $z=5.00$ as our lower redshift limit due to a combination of garnishing a large sample, adequately spanning physical properties (e.g. luminosity, age of the Universe) and to highlight the parts of $L-z$ parameter space where $z>5$ quasars still wait to be discovered.

This paper is organized as follows.  In Section 2, we present the assembled list of the 463 $z\geq5.00$ VH$z$Qs that we have compiled. We then give a high-level overview of the photometric surveys and datasets we use and present the photometry of the VH$z$Qs. In Section 3 we ... and In Section 4 we ... We conclude in Section 5 and present all the necessary details to obtain our dataset in the Appendices.

We present all our photometry and magnitudes on the AB zero-point system \citep{Oke_Gunn1983, Fukugita1996}.  This includes the near-infrared, as well as the mid-infrared magnitudes. Appendix~\ref{sec:filters} gives the AB to Vega transforms for a wide range of optical, NIR and MIR filters. We use a flat $\Lambda$CDM cosmology with $H0 = 67.7$ km s$-1$ Mpc$−1$, $\Omega_{\rm M} = 0.307$, and $\Omega_{\Lambda} = 0.693$ (Planck Collaboration et al. 2016) to be consistent with \citet{Banados2016} and all logarithms are to the base 10.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%  S E C T I O  N   2         S E C T I O  N   2           S E C T I O  N   2       S E C T I O  N   2
%%  S E C T I O  N   2         S E C T I O  N   2           S E C T I O  N   2       S E C T I O  N   2
%%  S E C T I O  N   2         S E C T I O  N   2           S E C T I O  N   2       S E C T I O  N   2
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%%trim=l b r t
%  \includegraphics[width=18.6cm, clip,trim=14mm 4mm 10mm 10mm]
  \includegraphics[width=8.6cm, clip,trim=32mm 4mm 32mm 10mm]
  {/cos_pc19a_npr/programs/quasars/highest_z/SEDs/filters_vs_QSOstars_z7pnt0.pdf}
  \centering
  \vspace{-12pt}
  \caption[]
  {The spectral bands used by different survey telescopes and that are relevant here.
    The $grizy$ filters are from the Pan-STARRS survey. The $JHK$ are from 
    UKIRT/WFCAM, while $K_{s}$ is a VISTA/VIRCAM filter. The 
    The narrow W-band centered at $\lambda\approx14,500$\AA\ is a CFHT/Wircam filter. 
    [NJC: Do we use the W-band anywhere?]
    The WISE passbands  W1-4 are also presented.
    The quasar spectrum is a composite based on \citet{VdB2001} and 
    \citet{Banados2016}. The L and T dwarf spectra are from \citet{Cushing2006}. 
  }
  \label{fig:filters}
\end{figure}

\vspace{-16pt}
\section{Method and Data}
Quasars are generally identified by photometric selection followed by
spectroscopic confirmation. Here, we reverse this method obtaining
first a list of spectroscopic quasars and then obtain photometric
information.

We have compiled a list of 463 quasars with redshifts $z\geq5.00$. We use all the $z\geq5.00$ quasars that have been discovered, spectroscopically confirmed and published as of 2018-Dec-31. We then obtain optical, near-infrared and mid-infrared photometry for the spectral dataset. The optical data comes from the Panoramic Survey Telescope and Rapid Response System (Pan-STARRS) survey \citep{Chambers2016}. The near-infrared data comes from two sources: first, the WFCAM \citep[][]{Casali2007} on the UKIRT, primarly, but not exclusively, as part of the UKIRT Infrared Deep Sky Survey \citep[UKIDSS; ][]{Lawrence2007}.  And second, data from the VIRCAM on the VISTA \citep[][]{Emerson2006, Dalton2006}. The mid-infrared, $\lambda=3-30\mu$m wavelength data is from the the Wide-Field Infrared Survey Explorer \citep[WISE;][]{Wright2010, Cutri2013} mission. 
%%
For reference, Figure~\ref{fig:filters} displays the wavelength and normalised transmission of the filters in question.

\subsection{Spectroscopy} 
We compile the list of all known, spectroscopically confirmed
quasars from the literature.  This list was complied from a range of
surveys and papers.  Specifically, we use data from:
\citet{Banados2014, Banados2016, Banados2018}, 
\citet{Becker2015}, 
\citet{Calura2014}, 
\citet{Carilli2007, Carilli2010}, 
\citet{Carnall2015}, 
\citet{Cool2006}, 
\citet{DeRosa2011}, 
\citet{Fan2000, Fan2001c, Fan2003, Fan2004, Fan2006, Fan2018}, 
\citet{Goto2006}, 
\citet{Ikeda2017}, 
\citet{Jiang2008, Jiang2009, Jiang2015, Jiang2016},   
\citet{Kashikawa2015}, 
\citet{Koptelova2017}, 
\citet{Kim2015, Kim2018},  
\citet{Kurk2007, Kurk2009}, 
\citet{Leipski2014}, 
\citet{Mahabal2005}, 
\citet{Matsuoka2016,  Matsuoka2018a, Matsuoka2018b},   
\citet{Mazzucchelli2017}, 
\citet{Morganson2012}, 
\citet{Mortlock2009, Mortlock2011},
\citet{McGreer2006, McGreer2013},  
\citet{Reed2015, Reed2017}, 
\citet{Stern2007},  
\citet{Tang2017}, 
\citet{Venemans2007, Venemans2012, Venemans2013, Venemans2015a, Venemans2015b, Venemans2016},
\citet{WangF2016, WangF2017, WangF2018a, WangF2018b},
\citet{Willott2007, Willott2009, Willott2010a, Willott2013b, Willott2015}, 
\citet{Wu2015} 
\citet{YangJ2018a, YangJ2018b}  
and 
\citet{Zeimann2011}. 
%We have obtained a list of 463 spectroscopically confirmed quasars with redshifts $z\geq5.00$.

Most of these objects are easily identified by their broad Ly$\alpha$
emission line, \nv emission and characteristic shape blueward of
1215\AA\ in the rest-frame. As we shall see, some of the recently
discovered objects are close to the galaxy luminosity function
characteristic luminosity $M^{*}$, and some have relatively weak or
maybe even completely absorbed Ly$\alpha$ \citep[e.g. Figures 7 and 10
in][]{Banados2016}. We leave aside detailed investigation and
discussion into spectral features and line strengths, and take as
given the published spectra and redshift identifications.

The breakdown of how many VH$z$Q each survey reports is given in Table~\ref{tab:surveys}. The Sloan Digital Sky Survey (SDSS) and the Pan-STARRS1 (PS1; PSO in Table~\ref{tab:surveys}) survey and alone identified over half (54.6\%) of the VH$z$Q population. Data from the Hyper Suprime-Cam (HSC) on the Subaru telecope is responsible for 13.6\% of our dataset (HSC+SHELLQs in Table~\ref{tab:surveys}). The combination of surveys is also vital for identifying VH$z$Qs. The UKIDSS Large Area Survey (ULAS) on its own, or in combination with other surveys is responsible for 6.5\% of the sample (SUV+ULAS) including the highest-$z$ object. Where more than one survey is used for the high-redshift identification (e.g. via shorter-band veto and longer wavelength detection) we follow the discovery paper naming convention.

\begin{table}
\begin{tabular}{l r r l}
\hline  \hline
Survey              & \# VH$z$Qs & (\%) & Survey reference  \\
\hline  
  ATLAS             &     4    &   ( 0.86)    &  \citet{Shanks2015} \\
  CFHQS            &   20    &   ( 4.32)    &  \citet{Willott2007} \\
  DELS               &   16    &   ( 3.46)    &  \citet{Dey2018} \\
  ELAIS              &     1    &   ( 0.22)    &  \citet{Vaisanen2000} \\
  FIRST              &     1    &   ( 0.22)    &  \citet{Becker1995} \\
  HSC                 &    8    &   ( 1.73)    & \citet{Miyazaki2018} \\
  IMS                 &     5    &   ( 1.08)     &  \citet{Kim2015} \\
  MMT               &   12    &   ( 2.59)     &  \citet{McGreer2013} \\
  NDWFS           &     1    &   ( 0.22)    &  \citet{JD1999} \\
  PSO                 &   83   &   (17.93)   &   \citet{Kaiser2002, Kaiser2010} \\
  RD                   &     1   &   ( 0.22)    &  \citet{Mahabal2005} \\
  SDSS                &  170  &    (36.72)    & \citet{EDR} \\
 SDWISE$^{b}$    &   27    &  ( 5.83)    &   \citet{WangF2016} \\
%% SHELLQs    &  $^{c}$63  &  (13.6)   &  \citet{Matsuoka2016} \\  %% if you include 8 objects from the first HSC paper...
  SHELLQs         &    55    &   (11.88)  &  \citet{Matsuoka2016}     \\  
  SUV$^{c}$       &   20     &    ( 4.32)  & \citet{YangJ2017} \\
  UHS               &    1      &  ( 0.22)     &  \citet{WangF2017} \\
  ULAS               &   10   &   ( 2.16)     & \citet{Lawrence2007} \\
  VDES$^{d}$       &   17  &    ( 3.67)     &  \citet{Reed2017} \\
  VHS                 &     1  &     ( 0.22)    & \citet{WangF2018b} \\
  VIK                 &     9    &  ( 1.94)    &  \citet{Edge2013} \\
  VIMOS           &    1      &  ( 0.22)     &   \citet{LeFevre2003} \\
\hline  \hline
\end{tabular}
\caption{The source and number of the VH$z$Q, with the key survey reference also given. 
  Recent survey name and acronyms include: 
  $^{a}$DESI Legacy Imaging Survey; 
  $^{b}$SDWISE = SDSS+WISE; 
  $^{a}$SUV  = SDSS-ULAS/VHS; 
  $^{c}$VDES = VHS/VIKING+DES; 
%  $^{c}$Includes 8 objects with a HyperSuprimeCam \citep[HSC; ][]{Miyazaki2018} designation.
}
      \label{tab:surveys}
\end{table}

\begin{figure}
%%trim=l b r t
  \includegraphics[width=8.0cm, clip, trim=10mm 0mm 0mm 0mm]
  {/cos_pc19a_npr/programs/quasars/highest_z/Nofz/Nofz_0pnt075bins_20181211.png}
  \centering
  \vspace{-12pt}
  \caption[]
  {The redshift distribution $N(z)$ of the VH$z$Q sample. 
    The bins are $\delta-z=0.075$ in width. }
  \label{fig:Nofz}
\end{figure}

The redshifts for the VH$z$Qs generally come from the measurement of
broad UV/optical emission lines. There are far infra-red emission
lines e.g. \cii~158 micron available for several objects, but at the
level of our current analysis broadline redshifts are sufficient.

\begin{table}
\centering
\begin{tabular}{l r  r}
\hline \hline
$z \geq$  & Age / Myr & No. of objects \\
\hline 
7.50         &    700         &   1   \\
7.00         &    767         &   4   \\
6.78         &    800         &  14   \\
6.50         &    845         &  40   \\
6.19         &   900          &  86   \\
6.00         &   937          &  170   \\
5.70         & 1000          &  267   \\
5.00         & 1180          &  463   \\
\hline \hline
\end{tabular}
\caption{The number of objects at or above a given redshift. 
The age of the Universe in megayears is also given. }
      \label{tab:ages}
\end{table}

The number of objects at or above various redshifts, along with the 
corresponding age of the Universe is given in Table~\ref{tab:ages}.

The $N(z)$ redshift histogram is given for the sample in Figure~\ref{fig:Nofz}. 
We split the contribution up by survey. For clarity we show the individual 
surveys of SDSS, PS1, HSC, the ULAS detection, and tally the remaining 
surveys together (``various''). In Section~\ref{sec:Lz} we discuss the form 
of the $N(z)$ as well as the coverage of the luminosity-redshift $L-z$ plane.

\subsection{Optical Photometry}
We query the Panoramic Survey Telescope and Rapid Response System (Pan-STARRS)\footnote{\href{https://outerspace.stsci.edu/display/PANSTARRS}{https://outerspace.stsci.edu/display/PANSTARRS}} Data Release 1 (DR1) Catalog Archive Server Jobs System (CasJobs) service at \href{http://mastweb.stsci.edu/ps1casjobs/}{mastweb.stsci.edu/ps1casjobs/}. The Pan-STARRS1 (PS1) survey observed the 30,000 deg$^{2}$ of sky north of declination $\delta = -30$ degrees in the five $grizy$ filters. PS1 is the first part of Pan-STARRS to be completed and is the basis for the DR1. \citet{Chambers2016}, \citet{Magnier2016a}, \citet{Waters2016}, \citet{Magnier2016b}, \citet{Magnier2016c} and \citet{Flewelling2016} describe the instrument, survey, and data analyses. The principal science product of the PS1 survey is the catalog accessible through the CasJobs interface.

The PS1 Casjobs SQL Server is located at
\href{http://mastweb.stsci.edu/ps1casjobs}{mastweb.stsci.edu/ps1casjobs}.
The top level documentation is given
\href{https://outerspace.stsci.edu/display/PANSTARRS/PS1+Source+extraction+and+catalogs}{here}
while the description of tables is given
\href{https://outerspace.stsci.edu/display/PANSTARRS/PS1+Source+extraction+and+catalogs#PS1Sourceextractionandcatalogs}{here}.

The main tables are the
\href{https://outerspace.stsci.edu/display/PANSTARRS/PS1+ObjectThin+table+fields}{{\tt
objectThin}} and
\href{https://outerspace.stsci.edu/display/PANSTARRS/PS1+MeanObject+table+fields}{{\tt
meanObject}} tables.  We query and return the mean PSF magnitudes from
the $grizy$ filters ({\tt MeanPSFMag}) which are in the AB system for
our 463 VH$z$Q sample. Our SQL and links to the main PS1 DR1 tables
are given at  Appendix~\ref{sec:PS1_SQL}.

\subsection{Near-infrared photometry}~\label{sec:NIR_data} 
The near-infrared data in this paper comes from the Wide Field
Astronomy Unit's \href{https://www.roe.ac.uk/ifa/wfau/}{(WFAU)}
Science Archives for UKIRT-WFCAM, the WFCAM Science Archive
\citep[WSA; ][]{Hambly2008} and VISTA-VIRCAM, the VISTA Science
Archive \citep[VSA; ][]{Cross2012}. These archives were developed for
the VISTA Data Flow System \citep[VDFS][]{VDFS}.

We access both the WSA and the VSA and include all non-proprietary
WFCAM data, which covers all public surveys and PI projects from
Semester 05A to 2017-Jan-01, and all non-proprietary VISTA data, which
covers all public surveys and PI projects from science verification on
2009-Oct-15 to 2016-Apr-01.

Here we are not just quering the WSA or VSA data tables. We are taking
a list of objects (positions) are performing matched aperature
(``forced'') photometry on the NIR imaging data. As such, we generate
a set of tables that are different in subtle ways to the regular
``Detection'' tables.  The two most important tables for our needs are
the {\tt [w/v]serv1000MapRemeasurement} and {\tt
[w/v]serv1000MapRemeasAver}.

We produce and provide a two new databases with all the necessary
quantaties and measurements to fully reproduce our tables, figures and
results herein. Morever, these databases report considerably more
information than we report here. Full documentation can be found at
the \href{http://wsa.roe.ac.uk/www/wsa_browser.html}{WSA Schema
Browser} and the
\href{http://horus.roe.ac.uk/vsa/www/vsa_browser.html}{VSA Schema
Browser}.

  \subsubsection{Averaging matched photometry}
  The data was processed using a matched-aperture photometry method
  where flux is measured at the spectroscopic position of the quasar,
  without necessarily knowing if there is a formal detection in the NIR
  photometry beforehand. The matched-aperture pipeline is discussed in
  \citet{Cross2013} and with fuller details to appear in a forthcoming
  paper (Cross et al., 2019, in prep).

  We query the WSA and VSA performing matched-aperature photometry at
  the positions of our 463 VH$z$Qs. This database is world-readable and
  we give the full receipe and relevant SQL queries for accessing both
  databases in Appendix~\ref{sec:SQL} as well as online.

  The photometry in a single epoch image often has low
  signal-to-noise.  The advantage of matched aperture photometry on
  quasars is that co-adding is relatively simple if each epoch is taken
  in the same aperture and the aperture photometry has been corrected to
  total. Indeed, the standard aperture corrections work well for point
  sources. Coadding using the matched-aperature photometry, where the
  individual epochs are taken from multiple projects with different
  pointings and orientations, should help with issues such as scattered
  light, pixel distortion and aperture corrections.

  We average the aperture corrected calibrated fluxes (e.g. {\bf
  aperJky3}), and then convert to magnitudes. Since we do not have a
  deep image for each set of averages, we cannot calculate non-aperture
  corrected values, so the photometry is only appropriate for
  point-sources.

  \begin{equation}
    \bar{F} = \frac{\sum_i^N (w_i\,F_i)}{\sum_i^N w_i}  
    \label{eq:avg}
  \end{equation}
  where $F_i$ is the $i^{th}$ epoch measurement of a parameter to be
  averaged such as the aperture corrected calibrated flux in a $1\arcsec$ aperture
  ({\bf aperJky3}) and $\bar{F}$ is the weighted mean average of this parameter.
  The weight for each epoch $w_i=1/(\sigma_{F})^2$ if the epoch is included and 
  $w_i=0$ if an epoch is excluded for quality control purposes.

  We calculate a set of averaged catalogues, for each pointing and filter, based
  on the requirements in \verb+RequiredMapAverages+, in these cases over time
  spans of 7, 14, 30, 91, 183 days, 365 days, 730 days, over 10 epochs and
  over all epochs. The averaging process starts at the first epoch and works onwards 
  from there. Again, we present these measurements in the new SQL tables.

  We detect 304 unique quasars in the WFCAM WSA database, 203 quasars are 
  detected in the VISTA VSA database with 114 objects in common with both
  WFCAM and VISTA data.  We give the necessary SQL queries syntax in Appendix~\ref{sec:SQL}.

\subsection{MIR data}
The MIR data for this study comes from the Wide-field Infrared Survey
Explorer (WISE) mission, and we utlize data from the WISE cryogenic
and NEOWISE (Mainzer et al. 2011 ApJ, 731, 53) post-cryogenic survey
phases.

We use data from the the beginning of the WISE mission \citep[2010 January; ][]{Wright2010} through the fourth-year of NEOWISE-R operations \citep[2017 December;]{Mainzer2011}.  More specifically, we utilise the recently released ``unWISE Catalog'' \citet{MeisnerSchlafly2019}. The unWISE effort\footnote{http://unwise.me/} is the unblurred coadds of the WISE imaging using the AllWISE and NEOWISE-R stacked data \citep{Lang2014, Meisner2018a, Meisner2018b}. %{\bf Few more words here on the unWISE catalog...}.

All fluxes in the unWISE catalog are reported there are in ``Vega nanoMaggies'', with the Vega magnitude of a source is given by $m_{\rm Vega} = 22.5−2.5\log(f)$, where $f$ is the source flux. The absolute calibration for unWISE is ultimately inherited from AllWISE through the calibration of \citet{Meisner2017}. This inheritance depends on details of the PSF normalization at large radii, which is uncertain. Subtracting 4 mmag from the unWISE W1, and 32 mmag from unWISE W2 fluxes improves the agreement between unWISE and AllWISE fluxes.

Thus to convert unWISE Vega magnitudes onto the AB system, we have    
\begin{eqnarray*}
        {\rm W1}_{\rm AB, unWISE}  & = &   22.5 - 2.5 \log(f_{\rm W1}) - 0.004 + 2.699 \\
        {\rm W2}_{\rm AB, unWISE}  &  = &  22.5 - 2.5 \log(f_{\rm W2}) - 0.032 + 3.339 
\end{eqnarray*}
%Henceforth, when talking about we will simply refer to ``W1'' and ``W2'' as shorthand for ${\rm W1}_{\rm AB, unWISE}$ and {\rm W2}_{\rm AB, unWISE}
those two numbers are how far our flux measurements differed on average from the AllWISE flux estimates at high Galactic latitudes.  These tweaks will move you closer to the AllWISE absolute calibration, which we haven't tried to improve upon.
%We note though that 
%Especially in W1, though, this correction is uncertain, because the average offset between AllWISE and unWISE depends on magnitude, and we don't really know
%- what the source of the trend is,
%- which of unWISE and AllWISE is correct, or
%- where the AllWISE tie to AB/Vega was done (though presumably it was done on the bright end?).

For the our MIR variability investigations, we do not use the unWISE coadds, but instead use 
the \href{http://wise2.ipac.caltech.edu/docs/release/allwise/}{AllWISE} catalogue and the \href{http://wise2.ipac.caltech.edu/docs/release/neowise/neowise_2018_release_intro.html}{NEOWISE 2018 Data Release}. NEOWISE 2018 makes available the 3.4 and 4.6 μm (W1 and W2) single-exposure images and extracted source information that was acquired between 2016 December 13 and 2017 December 13 UTC, which was the fourth year of survey operations of the Near-Earth Object Wide-field Infrared Survey Explorer Reactivation Mission (NEOWISE; Mainzer et al. 2014, ApJ, 792, 30). The fourth year NEOWISE data products are concatenated with those from the first three years into a single archive.

The WISE scan pattern leads to coverage of the full-sky approximately once every six months (a``sky pass''), but the satellite was placed in hibernation in 2011 February and then reactivated in 2013 October. Hence, our light curves have a cadence of 6 months with a 32 month sampling gap.

Tables~\ref{tab:THE_TABLE}, \ref{tab:THE_TABLE_NIR} and~\ref{tab:THE_TABLE_MIR}, represent the culmination of this effort, and we now describe the assembly of their contents in more detail.

\input{table_tex/THE_TABLE_v0pnt96_TOP10_betterformat}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%  S E C T I O  N   3         S E C T I O  N   3           S E C T I O  N   3       S E C T I O  N   3
%%  S E C T I O  N   3         S E C T I O  N   3           S E C T I O  N   3       S E C T I O  N   3
%%  S E C T I O  N   3         S E C T I O  N   3           S E C T I O  N   3       S E C T I O  N   3
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
Having collated the sample of 463 VH$z$Qs, and obtained their optical,
near- and mid-infrared photometry we report here the various
photometric properites of the quasars.

First, we will concentrate on detection rate in the infrared, go on to
report on the color-redshift and color-color properties of our sample
and then report on how the current sample populates the
luminosity-redshift $Lz$-plane.

    %\subsection{Detection Rates in the optical}

    \subsection{Detection Rates in the NIR}
    Table~\ref{tab:nir_detection} gives the detection rates for the 
    VH$z$Qs in the NIR $YJHK/K_{s}$-bands. 
    The first thing to note is that the coverage of the NIR surveys 
    for example from the UKIDSS LAS and VISTA VHS, does
    not overlap the full area for where the VH$z$Qs are detected.

    \begin{table}
          \centering

      \begin{tabular}{l r l}
        \hline  \hline
        Selection   & number detected (\%) \\
        \hline  
        Any band ($ZYJHK/K_{s}$   &  394  (92.9) \\
        $Z$-band    &  72  (17.0) \\
        $Y$-band    &  249  (58.7) \\
        $J$-band    &  391  (92.2) \\
        $H$-band    &  258  (60.8) \\
        $K$ or $Ks$-band    &  297  ( ) \\
        \hline  \hline
      \end{tabular}
      \caption{Detection rate of VH$z$Qs in the near-infrared.}
      \label{tab:nir_detection}
    \end{table}

     There are 17 objects that have no NIR detections. 
     2 of these (PSOJ053.9605-15.7956 and PSO J056.7168-16.4769) 
     have been observed, but are out of our queried time range. 
     10 objects have not been observed (or at least the data is not in the WSA/VSA 
     archives yet) and 5 objects at $\geq +60$ deg$^{2}$ are too far north for UKIRT 
     and cannot be observed.

>>>>

\subsection{The L-$z$ Plane}
Having obtained an as-near-to-homogenous set of photometry as we can, 
we are now in a position to calculate the Absolute Magnitudes of the VH$z$Q 
sample and in particulare the absolute magnitude at rest-frame 1450\AA\ , $M_{1450}$, 
which is a key physical quantity and goes directly towards the quasar luminosity 
function and thus the reionization of hydrogen calculation.
